---
title: "mathys_DEG_analysis_subcluster_MAST_interaction"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This analysis was orignially done in the mathys_DEG_analysis_subcluster_sex-cogdx_apoe-cogdx.Rmd
```{r}
library(ggpubr)
library(ggplot2) 
library(ggrepel)
library(dplyr)
library(EnhancedVolcano)
library(clusterProfiler)
library(org.Hs.eg.db)
# library(AnnotationHub)
library(patchwork)
library(scales)
library(VennDiagram)
library(data.table)
library(RColorBrewer)
library(stringr)
library(ReactomePA)
library(ComplexHeatmap)
library(colorRamp2)
library(reshape2)
library(ggdendro)
library(tidyr)
library(grid)
library(GOtest)
library(circlize)

source("/sc/arion/work/yug03/functions.R")
```


```{r}
processed.path <- "/sc/arion/projects/zhangb03a/shared/ROSMAP/Synapse/snRNAseq_MIT/GeneExpression/10x/processed/"
all.meta.data <- readRDS("/sc/arion/projects/zhangb03a/yug03/mathys/snRNA/merged_all_celltype_meta.RDS")
all.meta.data

processed.ls <- c("Astrocytes", "Excitatory_neurons_set1", "Excitatory_neurons_set2", "Excitatory_neurons_set3", "Inhibitory_neurons", "Oligodendrocytes", "OPCs", "Immune_cells", "Vasculature_cells")

mathys_annot_table <- read.table("/sc/arion/projects/zhangb03a/shared/ROSMAP/Synapse/snRNAseq_MIT/GeneExpression/10x/processed/Raw/genes.tsv", sep="\t", header = T)

celltype.df <- read.csv("/sc/arion/work/yug03/mathys/snRNA/celltype_df.txt", sep="\t")
celltype.df
```
```{r}
length(unique(all.meta.data$projid))

traits <- read.csv("/sc/arion/projects/zhangb03a/shared/ROSMAP/Synapse/Metadata/dataset_707_basic_02-08-2022.clean.txt", sep="\t")
rownames(traits) <- traits$projid
traits$sex <- ifelse(traits$msex == 0, "F", "M")
# traits$apoe2 <- ifelse(traits$apoe_genotype == "24", "4x", ifelse(traits$apoe_genotype == "34", "4x", ifelse(traits$apoe_genotype == "44", "4x", ifelse(traits$apoe_genotype == "22", "2x", ifelse(traits$apoe_genotype == "23", "2x", traits$apoe_genotype)))))
traits$apoe2 <- ifelse(traits$apoe_genotype == "24", "24", ifelse(traits$apoe_genotype == "34", "4x", ifelse(traits$apoe_genotype == "44", "4x", ifelse(traits$apoe_genotype == "22", "2x", ifelse(traits$apoe_genotype == "23", "2x", traits$apoe_genotype)))))
traits$apoe2 <- paste0("e", traits$apoe2)
traits$cogdx2 <- ifelse(traits$cogdx == 1, "NCI", ifelse(traits$cogdx == 2, "MCI", ifelse(traits$cogdx == "4", "AD", traits$cogdx)))
# traits$cogdx2 <- factor(traits$cogdx2 , levels = c("NCI", "MCI", "AD"))
traits$sex_apoe2 <- paste(traits$sex, traits$apoe2, sep = '_') #sample meta data
traits$sex_apoe2_cogdx <- paste(traits$sex, traits$apoe2, traits$cogdx2, sep = '_')
traits$sex_cogdx2 <- paste(traits$sex, traits$cogdx2, sep = '_') #sample meta data
traits$apoe2_cogdx2 <- paste(traits$apoe2, traits$cogdx2, sep = '_') #sample meta data

all.meta.data <- left_join(all.meta.data,traits[, c("projid","sex","apoe2","cogdx2","sex_apoe2", "sex_apoe2_cogdx", "sex_cogdx2", "apoe2_cogdx2", "pmi", "age_death")],by="projid")
all.meta.data
```




# load original data for plotting
```{r}
ori.ls <- list()
for (cellfile in unique(celltype.df$processed.ls3)[c(1,2,4)]){
# for (cellfile in unique(celltype.df$processed.ls3)[c(2,6,8)]){
  print(cellfile)
  pbmc.ori <- readRDS(paste0(processed.path, cellfile, ".rds"))
  print(dim(pbmc.ori[["RNA"]]@data))
  print(table(pbmc.ori@meta.data$cell_type_high_resolution))
  print(paste("Num total sample: ", length(unique(pbmc.ori@meta.data$projid))))
  
  pbmc <- CreateSeuratObject(pbmc.ori[["RNA"]]@counts, meta.data = pbmc.ori@meta.data)
  print(dim(pbmc[["RNA"]]@counts))
  print(table(pbmc@meta.data$cell_type_high_resolution))
  print(paste("Num total sample: ", length(unique(pbmc@meta.data$projid))))
  
  # normalize 
  pbmc <- NormalizeData(pbmc, normalization.method = "LogNormalize", scale.factor = 10000)
  print(rowSums(pbmc[["RNA"]]@data)[1:5])
  
  print(unique(as.character(pbmc@meta.data$cell_type_high_resolution)))
  pbmc@meta.data$celltype <- gsub( " .*$", "", pbmc@meta.data$cell_type_high_resolution)
  print(unique(pbmc@meta.data$celltype))
  
  Idents(pbmc) <- "cell_type_high_resolution"
  
  print(table(Idents(pbmc)))
  print(head(pbmc@meta.data))
  
  df <- left_join(pbmc@meta.data,traits[, c("projid","sex","apoe2","cogdx2","sex_apoe2", "sex_apoe2_cogdx", "sex_cogdx2", "apoe2_cogdx2", "pmi", "age_death")],by="projid")
  rownames(df) <- rownames(pbmc@meta.data)
  pbmc <- AddMetaData(pbmc, df)
  # print(head(pbmc@meta.data))
  
  
  print(table(pbmc@meta.data$sex_apoe2_cogdx))
  
  print(paste("Num  sample NA for pmi:",  length(unique(pbmc@meta.data[is.na(pbmc@meta.data$pmi), "projid"]))))
  print(paste("Num  cell NA for pmi:",  sum(is.na(pbmc@meta.data$pmi))))
  print(paste("Sample details NA for pmi:",  unique(pbmc@meta.data[is.na(pbmc@meta.data$pmi), "sex_apoe2_cogdx"])))
  
  # pbmc.sub <- pbmc[, pbmc$cogdx2 %in% c("AD", "MCI", "NCI") & !is.na(pbmc$pmi) & pbmc$apoe2 != "eNA" & pbmc$apoe2 != "e24"]
  pbmc.sub <- pbmc[, pbmc$cogdx2 %in% c("AD", "NCI") & !is.na(pbmc$pmi) & pbmc$apoe2 != "eNA" & pbmc$apoe2 != "e24"]
  
  print(table(pbmc.sub@meta.data$sex_apoe2_cogdx))
  print(table(pbmc.sub@meta.data$sex_cogdx))
  print(table(pbmc.sub@meta.data$apoe2_cogdx2))
  print(dim(pbmc.sub[["RNA"]]@data))
  print(length(unique(pbmc.sub@meta.data$projid)))
  print(head(pbmc.sub@meta.data))
  
  ori.ls[[cellfile]] <- pbmc.sub
}
```


## Test MAST

```{r}
library(MAST)
```

```{r my_MASTDETest and my_IdentsToCells}
my_MASTDETest <- function(
  data.use,
  cells.1,
  cells.2,
  latent.vars = NULL,
  verbose = TRUE,
  ...
) {
  # Check for MAST
  if (!PackageCheck('MAST', error = FALSE)) {
    stop("Please install MAST - learn more at https://github.com/RGLab/MAST")
  }
  group.info <- data.frame(row.names = c(cells.1, cells.2))
  latent.vars <- latent.vars %||% group.info
  group.info[cells.1, "group"] <- "Group1"
  group.info[cells.2, "group"] <- "Group2"
  group.info[, "group"] <- factor(x = group.info[, "group"])
  latent.vars.names <- c("condition", colnames(x = latent.vars))
  latent.vars <- cbind(latent.vars, group.info)
  latent.vars$wellKey <- rownames(x = latent.vars)
  fdat <- data.frame(rownames(x = data.use))
  colnames(x = fdat)[1] <- "primerid"
  rownames(x = fdat) <- fdat[, 1]
  sca <- MAST::FromMatrix(
    exprsArray = as.matrix(x = data.use),
    check_sanity = FALSE,
    cData = latent.vars,
    fData = fdat
  )
  cond <- factor(x = SummarizedExperiment::colData(sca)$group)
  cond <- relevel(x = cond, ref = "Group1")
  SummarizedExperiment::colData(sca)$condition <- cond
  fmla <- as.formula(
    object = paste0(" ~ ", paste(latent.vars.names, collapse = "+"))
  )
  zlmCond <- MAST::zlm(formula = fmla, sca = sca)
  summaryCond <- MAST::summary(object = zlmCond, doLRT = 'conditionGroup2')
  summaryDt <- summaryCond$datatable
  summaryDt <- as.data.frame(summaryDt)
  p_val <- summaryDt[summaryDt[, "component"] == "H", 4]
  genes.return <- summaryDt[summaryDt[, "component"] == "H", 1]
  to.return <- data.frame(p_val, row.names = genes.return)
  return(to.return)
}

my_IdentsToCells <- function(
  object,
  ident.1,
  ident.2,
  cellnames.use
) {
  #
  if (is.null(x = ident.1)) {
    stop("Please provide ident.1")
  } else if ((length(x = ident.1) == 1 && ident.1[1] == 'clustertree') || is(object = ident.1, class2 = 'phylo')) {
    if (is.null(x = ident.2)) {
      stop("Please pass a node to 'ident.2' to run FindMarkers on a tree")
    }
    tree <- if (is(object = ident.1, class2 = 'phylo')) {
      ident.1
    } else {
      Tool(object = object, slot = 'BuildClusterTree')
    }
    if (is.null(x = tree)) {
      stop("Please run 'BuildClusterTree' or pass an object of class 'phylo' as 'ident.1'")
    }
    ident.1 <- tree$tip.label[GetLeftDescendants(tree = tree, node = ident.2)]
    ident.2 <- tree$tip.label[GetRightDescendants(tree = tree, node = ident.2)]
  }
  if (length(x = as.vector(x = ident.1)) > 1 &&
      any(as.character(x = ident.1) %in% cellnames.use)) {
    bad.cells <- cellnames.use[which(x = !as.character(x = ident.1) %in% cellnames.use)]
    if (length(x = bad.cells) > 0) {
      stop(paste0("The following cell names provided to ident.1 are not present in the object: ", paste(bad.cells, collapse = ", ")))
    }
  } else {
    ident.1 <- WhichCells(object = object, idents = ident.1)
  }
  # if NULL for ident.2, use all other cells
  if (length(x = as.vector(x = ident.2)) > 1 &&
      any(as.character(x = ident.2) %in% cellnames.use)) {
    bad.cells <- cellnames.use[which(!as.character(x = ident.2) %in% cellnames.use)]
    if (length(x = bad.cells) > 0) {
      stop(paste0("The following cell names provided to ident.2 are not present in the object: ", paste(bad.cells, collapse = ", ")))
    }
  } else {
    if (is.null(x = ident.2)) {
      ident.2 <- setdiff(x = cellnames.use, y = ident.1)
    } else {
      ident.2 <- WhichCells(object = object, idents = ident.2)
    }
  }
  return(list(cells.1 = ident.1, cells.2 = ident.2))
}
```

```{r}
pbmc <- ori.ls$Immune_cells
pbmc@meta.data
table(pbmc@meta.data$cell_type_high_resolution)
```


```{r}
r <- sex_cogdx2.res$sex_cogdx.F_NCI_vs_M_NCI[sex_cogdx2.res$sex_cogdx.F_NCI_vs_M_NCI$Celltype == "T cells", ]
r
r[r$p_val_adj < 0.05, ]
# r[r$p_val_adj < 0.05 & abs(r$avg_log2FC) > log2(1.3), ]
```

The FC is different because in Rstudio, it use seurat 4 and above use R command line Seurat (v5). Seurat v5 tends to have higher FC
```{r}
logfc.threshold <- log(1)
latent.vars <- c('nCount_RNA','pmi','age_death')
result1 <- FindMarkers(object = pbmc, slot = "data", assay = NULL, ident.1 = "F_NCI", ident.2 = "M_NCI", group.by = "sex_cogdx2", subset.ident = "T cells", logfc.threshold = logfc.threshold, test.use = 'MAST', latent.vars = latent.vars, min.pct = 0.1)
names(result1)[names(result1) == 'p_val_adj'] <- 'p_val_adj_bonferroni'
print(nrow(result1))
result1$p_val_adj <- p.adjust(result1$p_val, n=nrow(result1), method = "BH")
result1
result1[result1$p_val_adj < 0.05, ]
# result1[result1$p_val_adj < 0.05 & abs(result1$avg_log2FC) > log2(1.3),]
```



mimic FindMarkers.Seurat 
```{r}
pbmc.sub <- pbmc[, pbmc@meta.data$cell_type_high_resolution == "T cells"]
data.use.1 <- pbmc.sub[["RNA"]]
Idents(pbmc.sub) <- "sex_cogdx2"
latent.vars <- c('nCount_RNA','pmi','age_death')

cellnames.use.1 <- colnames(data.use.1)

cells_1 <- my_IdentsToCells(
    object = pbmc.sub,
    ident.1 = "F_NCI",
    ident.2 = "M_NCI",
    cellnames.use = cellnames.use.1
  )

cells_1 <- sapply(
  X = cells_1,
  FUN = intersect,
  y = cellnames.use.1,
  simplify = FALSE,
  USE.NAMES = TRUE
)

latent.vars.1 <- FetchData(
      object = pbmc.sub,
      vars = latent.vars,
      cells = c(cells_1$cells.1, cells_1$cells.2)
    )

# my_MASTDETest(data.use.1@data[, c(cells_1$cells.1, cells_1$cells.2)], cells.1 = cells_1$cells.1, cells.2 = cells_1$cells.2, latent.vars = latent.vars.1, norm.method = "LogNormalize")

```

```{r my_FindMarkers.Seurat}
my_FindMarkers.Seurat <- function(
  object,
  ident.1 = NULL,
  ident.2 = NULL,
  latent.vars = NULL,
  group.by = NULL,
  subset.ident = NULL,
  assay = NULL,
  reduction = NULL,
  ...
) {
  if (!is.null(x = group.by)) {
    if (!is.null(x = subset.ident)) {
      object <- subset(x = object, idents = subset.ident)
    }
    Idents(object = object) <- group.by
  }
  if (!is.null(x = assay) && !is.null(x = reduction)) {
    stop("Please only specify either assay or reduction.")
  }
  if (length(x = ident.1) == 0) {
    stop("At least 1 ident must be specified in `ident.1`")
  }

  # select which data to use
  if (is.null(x = reduction)) {
    assay <- assay %||% DefaultAssay(object = object)
    data.use <- object[[assay]]
    cellnames.use <-  colnames(x = data.use)
  } else {
    data.use <- object[[reduction]]
    cellnames.use <- rownames(x = data.use)
  }
  cells <- my_IdentsToCells(
    object = object,
    ident.1 = ident.1,
    ident.2 = ident.2,
    cellnames.use = cellnames.use
  )
  cells <- sapply(
    X = cells,
    FUN = intersect,
    y = cellnames.use,
    simplify = FALSE,
    USE.NAMES = TRUE
  )
  if (!all(vapply(X = cells, FUN = length, FUN.VALUE = integer(length = 1L)))) {
    abort(
      message = "Cells in one or both identity groups are not present in the data requested"
    )
  }

  # fetch latent.vars
  if (!is.null(x = latent.vars)) {
    latent.vars <- FetchData(
      object = object,
      vars = latent.vars,
      cells = c(cells$cells.1, cells$cells.2)
    )
  }

  # check normalization method
  norm.command <- paste0("NormalizeData.", assay)
  norm.method <- if (norm.command %in% Command(object = object) && is.null(x = reduction)) {
    Command(
      object = object,
      command = norm.command,
      value = "normalization.method"
    )
  } else if (length(x = intersect(x = c("FindIntegrationAnchors", "FindTransferAnchors"), y = Command(object = object)))) {
    command <- intersect(x = c("FindIntegrationAnchors", "FindTransferAnchors"), y = Command(object = object))[1]
    Command(
      object = object,
      command = command,
      value = "normalization.method"
      )
    } else {
    NULL
  }
  
  cat("data.use: ", identical(data.use.1, data.use), "\n")
  cat("latent.vars: ", identical(latent.vars.1, latent.vars), "\n")
  cat("cells$cells.1: ", identical(cells_1$cells.1, cells$cells.1), "\n")
  cat("cells$cells.2: ", identical(cells_1$cells.2, cells$cells.2), "\n")
  print(norm.method)
  
  # This actually calles FindMarkers.default!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
  print(class(data.use))
  getS3method("FindMarkers", class(data.use)[1])
  de.results <- my_unknown_function(
    object = data.use,
    latent.vars = latent.vars,
    cells.1 = cells$cells.1,
    cells.2 = cells$cells.2,
    norm.method = norm.method,
    ...
  )
  return(de.results)
}

my_FindMarkers.Seurat(pbmc[, pbmc@meta.data$cell_type_high_resolution == "T cells"],
  ident.1 = "F_NCI",
    ident.2 = "M_NCI",
  group.by = "sex_cogdx2",
  latent.vars = c('nCount_RNA','pmi','age_death'), min.pct = 0.1, slot = "data", logfc.threshold = 0, test.use = 'MAST')
```

wrong!
without using unknown_function and  FindMarkers.default
```{r}
logfc.threshold <- log(1)
latent.vars <- latent.vars.1
cells.1 <- cells_1$cells.1
cells.2 <- cells_1$cells.2

group.info <- data.frame(row.names = c(cells.1, cells.2))
latent.vars <- latent.vars %||% group.info
group.info[cells.1, "group"] <- "Group1"
group.info[cells.2, "group"] <- "Group2"
group.info[, "group"] <- factor(x = group.info[, "group"])
group.info

latent.vars.names <- c("condition", colnames(x = latent.vars))
latent.vars <- cbind(latent.vars, group.info)
latent.vars$wellKey <- rownames(x = latent.vars)
latent.vars

fdat <- data.frame(rownames(x = data.use.1@data))
colnames(x = fdat)[1] <- "primerid"
rownames(x = fdat) <- fdat[, 1]

data.use <- data.use.1@data[, rownames(latent.vars)]

sca <- MAST::FromMatrix(
  exprsArray = as.matrix(x = data.use),
  check_sanity = FALSE,
  cData = latent.vars,
  fData = fdat
)

cond <- factor(x = SummarizedExperiment::colData(sca)$group)
cond <- relevel(x = cond, ref = "Group1")
SummarizedExperiment::colData(sca)$condition <- cond
fmla <- as.formula(
  object = paste0(" ~ ", paste(latent.vars.names, collapse = "+"))
)
zlmCond <- MAST::zlm(formula = fmla, sca = sca)
summaryCond <- MAST::summary(object = zlmCond, doLRT = 'conditionGroup2')
summaryDt <- summaryCond$datatable
summaryDt <- as.data.frame(summaryDt)
p_val <- summaryDt[summaryDt[, "component"] == "H", 4]
genes.return <- summaryDt[summaryDt[, "component"] == "H", 1]
to.return <- data.frame(genes.return, p_val, row.names = genes.return)
to.return[order(to.return$p_val), ]

```

test unknown_function and FindMarkers.default 
```{r my_FindMarkers.default}
my_FindMarkers.default <- function(
  object,
  slot = "data",
  cells.1 = NULL,
  cells.2 = NULL,
  features = NULL,
  logfc.threshold = 0.1,
  test.use = "wilcox",
  min.pct = 0.01,
  min.diff.pct = -Inf,
  verbose = TRUE,
  only.pos = FALSE,
  max.cells.per.ident = Inf,
  random.seed = 1,
  latent.vars = NULL,
  min.cells.feature = 3,
  min.cells.group = 3,
  fc.results = NULL,
  densify = FALSE,
  ...
) {
  features <- features %||% rownames(x = object)
  # reset parameters so no feature filtering is performed
  if (test.use %in% DEmethods_noprefilter()) {
    features <- rownames(x = object)
    min.diff.pct <- -Inf
    logfc.threshold <- 0
  }
  # feature selection (based on percentages)
  alpha.min <- pmax(fc.results$pct.1, fc.results$pct.2)
  names(x = alpha.min) <- rownames(x = fc.results)
  features <- names(x = which(x = alpha.min >= min.pct))
  if (length(x = features) == 0) {
    warning("No features pass min.pct threshold; returning empty data.frame")
    return(fc.results[features, ])
  }
  alpha.diff <- alpha.min - pmin(fc.results$pct.1, fc.results$pct.2)
  features <- names(
    x = which(x = alpha.min >= min.pct & alpha.diff >= min.diff.pct)
  )
  if (length(x = features) == 0) {
    warning("No features pass min.diff.pct threshold; returning empty data.frame")
    return(fc.results[features, ])
  }

  # feature selection (based on logFC)
  if (slot != "scale.data") {
    total.diff <- fc.results[, 1] #first column is logFC
    names(total.diff) <- rownames(fc.results)
    features.diff <- if (only.pos) {
      names(x = which(x = total.diff >= logfc.threshold))
    } else {
      names(x = which(x = abs(x = total.diff) >= logfc.threshold))
    }
    features <- intersect(x = features, y = features.diff)
    if (length(x = features) == 0) {
      warning("No features pass logfc.threshold threshold; returning empty data.frame")
      return(fc.results[features, ])
    }
  }

  # subsample cell groups if they are too large
  if (max.cells.per.ident < Inf) {
    set.seed(seed = random.seed)
    if (length(x = cells.1) > max.cells.per.ident) {
      cells.1 <- sample(x = cells.1, size = max.cells.per.ident)
    }
    if (length(x = cells.2) > max.cells.per.ident) {
      cells.2 <- sample(x = cells.2, size = max.cells.per.ident)
    }
    if (!is.null(x = latent.vars)) {
      latent.vars <- latent.vars[c(cells.1, cells.2), , drop = FALSE]
    }
  }
  if (inherits(x = object, what = "IterableMatrix")){
    if(test.use != "wilcox"){
      stop("Differential expression with BPCells currently only supports the 'wilcox' method.",
           " Please rerun with test.use = 'wilcox'")
    }
    data.use <- object[features, c(cells.1, cells.2), drop = FALSE]
    groups <- c(rep("foreground", length(cells.1)), rep("background", length(cells.2)))
    de.results <- suppressMessages(
      BPCells::marker_features(data.use, group = groups, method = "wilcoxon")
    )
    de.results <- subset(de.results, de.results$foreground == "foreground")
    de.results <- data.frame(feature = de.results$feature,
                             p_val = de.results$p_val_raw)
    rownames(de.results) <- de.results$feature
    de.results$feature <- NULL
  } else {
    de.results <- my_PerformDE(
      object = object,
      cells.1 = cells.1,
      cells.2 = cells.2,
      features = features,
      test.use = test.use,
      verbose = verbose,
      min.cells.feature = min.cells.feature,
      latent.vars = latent.vars,
      densify = densify,
      ...
    )
  }
  de.results <- cbind(de.results, fc.results[rownames(x = de.results), , drop = FALSE])
  if (only.pos) {
    de.results <- de.results[de.results[, 2] > 0, , drop = FALSE]
  }
  if (test.use %in% DEmethods_nocorrect()) {
    de.results <- de.results[order(-de.results$power, -de.results[, 1]), ]
  } else {
    de.results <- de.results[order(de.results$p_val, -abs(de.results$pct.1-de.results$pct.2)), ]
    de.results$p_val_adj = p.adjust(
      p = de.results$p_val,
      method = "bonferroni",
      n = nrow(x = object)
    )
  }
  return(de.results)
}
DEmethods_noprefilter <- function() {
  c("DESeq2")
}
DEmethods_counts <- function() {
  c("negbinom", "poisson", "DESeq2")
}
DEmethods_latent <- function() {
  c('negbinom', 'poisson', 'MAST', "LR")
}
DEmethods_checkdots <- function() {
  c('wilcox', 'wilcox_limma', 'MAST', 'DESeq2')
}
DEmethods_nocorrect <- function() {
  c('roc')
}
my_PerformDE <- function(
  object,
  cells.1,
  cells.2,
  features,
  test.use,
  verbose,
  min.cells.feature,
  latent.vars,
  densify,
  ...
) {
  if (!(test.use %in% DEmethods_latent()) && !is.null(x = latent.vars)) {
    warning(
      "'latent.vars' is only used for the following tests: ",
      paste(DEmethods_latent(), collapse=", "),
      call. = FALSE,
      immediate. = TRUE
    )
  }
  if (!test.use %in% DEmethods_checkdots()) {
    CheckDots(...)
  }
  data.use <- object[features, c(cells.1, cells.2), drop = FALSE]
  if (densify){
    data.use <- as.matrix(x = data.use)
  }
  de.results <- switch(
    EXPR = test.use,
    'wilcox' = WilcoxDETest(
      data.use = data.use,
      cells.1 = cells.1,
      cells.2 = cells.2,
      verbose = verbose,
      ...
    ),
    'wilcox_limma' = WilcoxDETest(
      data.use = data.use,
      cells.1 = cells.1,
      cells.2 = cells.2,
      verbose = verbose,
      limma = TRUE,
      ...
    ),
    'MAST' = my_MASTDETest(
      data.use = data.use,
      cells.1 = cells.1,
      cells.2 = cells.2,
      latent.vars = latent.vars,
      verbose = verbose,
      ...
    ),
    stop("Unknown test: ", test.use)
  )
  return(de.results)
}

my_unknown_function <- function (object, slot = "data", cells.1 = NULL, cells.2 = NULL, 
    features = NULL, logfc.threshold = 0.25, test.use = "wilcox", 
    min.pct = 0.1, min.diff.pct = -Inf, verbose = TRUE, only.pos = FALSE, 
    max.cells.per.ident = Inf, random.seed = 1, latent.vars = NULL, 
    min.cells.feature = 3, min.cells.group = 3, pseudocount.use = 1, 
    mean.fxn = NULL, fc.name = NULL, base = 2, densify = FALSE, 
    norm.method = NULL, ...) 
{
    pseudocount.use <- pseudocount.use %||% 1
    data.slot <- ifelse(test = test.use %in% DEmethods_counts(), 
        yes = "counts", no = slot)
    data.use <- GetAssayData(object = object, slot = data.slot)
    counts <- switch(EXPR = data.slot, scale.data = GetAssayData(object = object, 
        slot = "counts"), numeric())
    fc.results <- FoldChange(object = object, slot = data.slot, 
        cells.1 = cells.1, cells.2 = cells.2, features = features, 
        pseudocount.use = pseudocount.use, mean.fxn = mean.fxn, 
        fc.name = fc.name, base = base, norm.method = norm.method)
    
    cat("data.use: ", identical(data.use.1@data, data.use), "\n")
    cat("latent.vars: ", identical(latent.vars.1, latent.vars), "\n")
    cat("cells$cells.1: ", identical(cells_1$cells.1, cells.1), "\n")
    cat("cells$cells.2: ", identical(cells_1$cells.2, cells.2), "\n")
    print(class(data.use))
    print(pseudocount.use)
    print(features)
  
    de.results <- my_FindMarkers.default(object = data.use, slot = data.slot,
        counts = counts, cells.1 = cells.1, cells.2 = cells.2,
        features = features, logfc.threshold = logfc.threshold,
        test.use = test.use, min.pct = min.pct, min.diff.pct = min.diff.pct,
        verbose = verbose, only.pos = only.pos, max.cells.per.ident = max.cells.per.ident,
        random.seed = random.seed, latent.vars = latent.vars,
        min.cells.feature = min.cells.feature, min.cells.group = min.cells.group,
        pseudocount.use = pseudocount.use, fc.results = fc.results,
        densify = densify, ...)
    return(de.results)
}

my_unknown_function(object = data.use.1,
    latent.vars = latent.vars.1,
    cells.1 = cells_1$cells.1,
    cells.2 = cells_1$cells.2,
    norm.method = "LogNormalize", test.use = "MAST", logfc.threshold = 0, min.pct = 0.1)


fc.results.1 <- FoldChange(object = data.use.1, slot = "data", 
        cells.1 = cells_1$cells.1, cells.2 = cells_1$cells.2, features = NULL, 
        pseudocount.use = 1, norm.method = "LogNormalize")
result2 <- my_FindMarkers.default(object = data.use.1@data, features = NULL,
    latent.vars = latent.vars.1,
    cells.1 = cells_1$cells.1,
    cells.2 = cells_1$cells.2,
    norm.method = "LogNormalize", test.use = "MAST", logfc.threshold = 0, min.pct = 0.1, pseudocount.use = 1, fc.results = fc.results.1)
result2

identical(result1$p_val, result2$p_val)
identical(result1$avg_log2FC, result2$avg_log2FC)
```

with FoldChange and FindMarkers.default
part 1 -still using PerformDE
```{r}
min.pct <- 0.1
min.diff.pct <- -Inf
slot <- "data"
only.pos <- FALSE
max.cells.per.ident <-  Inf
fc.results.1 <- FoldChange(object = data.use.1, slot = "data", 
        cells.1 = cells_1$cells.1, cells.2 = cells_1$cells.2, features = NULL, 
        pseudocount.use = 1, norm.method = "LogNormalize")

features <- NULL
features <- features %||% rownames(x = data.use.1@data)

# feature selection (based on percentages)
alpha.min <- pmax(fc.results.1$pct.1, fc.results.1$pct.2)
names(x = alpha.min) <- rownames(x = fc.results.1)
features <- names(x = which(x = alpha.min >= min.pct))
if (length(x = features) == 0) {
  warning("No features pass min.pct threshold; returning empty data.frame")
  return(fc.results.1[features, ])
}
alpha.diff <- alpha.min - pmin(fc.results.1$pct.1, fc.results.1$pct.2)
features <- names(
  x = which(x = alpha.min >= min.pct & alpha.diff >= min.diff.pct)
)
if (length(x = features) == 0) {
  warning("No features pass min.diff.pct threshold; returning empty data.frame")
  return(fc.results[features, ])
}

# feature selection (based on logFC)
if (slot != "scale.data") {
  total.diff <- fc.results.1[, 1] #first column is logFC
  names(total.diff) <- rownames(fc.results.1)
  features.diff <- if (only.pos) {
    names(x = which(x = total.diff >= logfc.threshold))
  } else {
    names(x = which(x = abs(x = total.diff) >= logfc.threshold))
  }
  features <- intersect(x = features, y = features.diff)
  if (length(x = features) == 0) {
    warning("No features pass logfc.threshold threshold; returning empty data.frame")
    return(fc.results.1[features, ])
  }
}

de.results <- my_PerformDE(
  object = data.use.1@data,
  cells.1 = cells_1$cells.1, cells.2 = cells_1$cells.2,
  features = features,
  test.use = "MAST",
  min.cells.feature = 3,
  latent.vars = latent.vars.1,
  densify = FALSE
)
de.results <- cbind(de.results, fc.results.1[rownames(x = de.results), , drop = FALSE])

de.results <- de.results[order(de.results$p_val, -abs(de.results$pct.1-de.results$pct.2)), ]
de.results$p_val_adj = p.adjust(
  p = de.results$p_val,
  method = "bonferroni",
  n = nrow(x = data.use.1@data)
)
de.results
identical(de.results, result2)
```
part 2 - using plain codes
```{r}
logfc.threshold <- log(1)
latent.vars <- latent.vars.1
cells.1 <- cells_1$cells.1
cells.2 <- cells_1$cells.2

data.use.2 <- data.use.1@data[features, c(cells.1, cells.2), drop = FALSE]

group.info <- data.frame(row.names = c(cells.1, cells.2))
latent.vars <- latent.vars %||% group.info
group.info[cells.1, "group"] <- "Group1"
group.info[cells.2, "group"] <- "Group2"
group.info[, "group"] <- factor(x = group.info[, "group"])
group.info

latent.vars.names <- c("condition", colnames(x = latent.vars))
latent.vars <- cbind(latent.vars, group.info)
latent.vars$wellKey <- rownames(x = latent.vars)
latent.vars

fdat <- data.frame(rownames(x = data.use.2))
colnames(x = fdat)[1] <- "primerid"
rownames(x = fdat) <- fdat[, 1]

sca <- MAST::FromMatrix(
  exprsArray = as.matrix(x = data.use.2),
  check_sanity = FALSE,
  cData = latent.vars,
  fData = fdat
)

cond <- factor(x = SummarizedExperiment::colData(sca)$group)
cond <- relevel(x = cond, ref = "Group1")
SummarizedExperiment::colData(sca)$condition <- cond
fmla <- as.formula(
  object = paste0(" ~ ", paste(latent.vars.names, collapse = "+"))
)
zlmCond <- MAST::zlm(formula = fmla, sca = sca)
summaryCond <- MAST::summary(object = zlmCond, doLRT = 'conditionGroup2')
summaryDt <- summaryCond$datatable
summaryDt <- as.data.frame(summaryDt)
p_val <- summaryDt[summaryDt[, "component"] == "H", 4]
genes.return <- summaryDt[summaryDt[, "component"] == "H", 1]
to.return <- data.frame(p_val, row.names = genes.return)

de.result <- cbind(to.return, fc.results.1[rownames(x = to.return), , drop = FALSE])

de.result <- de.result[order(de.result$p_val, -abs(de.result$pct.1-de.result$pct.2)), ]
de.result$p_val_adj = p.adjust(
  p = de.result$p_val,
  # method = "bonferroni",
  # n = nrow(x = data.use.1@data)
  method = "BH",
  n = nrow(x = de.result)
)
de.result
identical(de.result$p_val_adj, result1$p_val_adj)
```


summarize:
FindMarkers.Seurat -> unknown_function -> FoldChange -> FindMarkers.default -> PerformDE -> MASTDETest



### All in one

```{r}
pbmc <- ori.ls$Immune_cells
pbmc.sub <- pbmc[, pbmc@meta.data$cell_type_high_resolution == "T cells"]
data.use.1 <- pbmc.sub[["RNA"]]
Idents(pbmc.sub) <- "sex_cogdx2"
latent.vars <- c('nCount_RNA','pmi','age_death')

cellnames.use.1 <- colnames(data.use.1)

cells_1 <- my_IdentsToCells(object = pbmc.sub,ident.1 = "F_NCI",ident.2 = "M_NCI",cellnames.use = cellnames.use.1)

cells_1 <- sapply(X = cells_1,FUN = intersect,y = cellnames.use.1,simplify = FALSE,USE.NAMES = TRUE)

latent.vars.1 <- FetchData(object = pbmc.sub,vars = latent.vars,cells = c(cells_1$cells.1, cells_1$cells.2))

min.pct <- 0.1
logfc.threshold <- log(1)
min.diff.pct <- -Inf
slot <- "data"
only.pos <- FALSE
max.cells.per.ident <-  Inf
fc.results.1 <- FoldChange(object = data.use.1, slot = "data", 
        cells.1 = cells_1$cells.1, cells.2 = cells_1$cells.2, features = NULL, 
        pseudocount.use = 1, norm.method = "LogNormalize")

features <- NULL
features <- features %||% rownames(x = data.use.1@data)

# feature selection (based on percentages)
alpha.min <- pmax(fc.results.1$pct.1, fc.results.1$pct.2)
names(x = alpha.min) <- rownames(x = fc.results.1)
features <- names(x = which(x = alpha.min >= min.pct))
if (length(x = features) == 0) {
  warning("No features pass min.pct threshold; returning empty data.frame")
  return(fc.results.1[features, ])
}
alpha.diff <- alpha.min - pmin(fc.results.1$pct.1, fc.results.1$pct.2)
features <- names(
  x = which(x = alpha.min >= min.pct & alpha.diff >= min.diff.pct)
)
if (length(x = features) == 0) {
  warning("No features pass min.diff.pct threshold; returning empty data.frame")
  return(fc.results[features, ])
}

# feature selection (based on logFC)
if (slot != "scale.data") {
  total.diff <- fc.results.1[, 1] #first column is logFC
  names(total.diff) <- rownames(fc.results.1)
  features.diff <- if (only.pos) {
    names(x = which(x = total.diff >= logfc.threshold))
  } else {
    names(x = which(x = abs(x = total.diff) >= logfc.threshold))
  }
  features <- intersect(x = features, y = features.diff)
  if (length(x = features) == 0) {
    warning("No features pass logfc.threshold threshold; returning empty data.frame")
    return(fc.results.1[features, ])
  }
}


logfc.threshold <- log(1)
latent.vars <- latent.vars.1
cells.1 <- cells_1$cells.1
cells.2 <- cells_1$cells.2

data.use.2 <- data.use.1@data[features, c(cells.1, cells.2), drop = FALSE]

group.info <- data.frame(row.names = c(cells.1, cells.2))
latent.vars <- latent.vars %||% group.info
group.info[cells.1, "group"] <- "Group1"
group.info[cells.2, "group"] <- "Group2"
group.info[, "group"] <- factor(x = group.info[, "group"])
group.info

latent.vars.names <- c("condition", colnames(x = latent.vars))
latent.vars <- cbind(latent.vars, group.info)
latent.vars$wellKey <- rownames(x = latent.vars)
latent.vars

fdat <- data.frame(rownames(x = data.use.2))
colnames(x = fdat)[1] <- "primerid"
rownames(x = fdat) <- fdat[, 1]

sca <- MAST::FromMatrix(
  exprsArray = as.matrix(x = data.use.2),
  check_sanity = FALSE,
  cData = latent.vars,
  fData = fdat
)

cond <- factor(x = SummarizedExperiment::colData(sca)$group)
cond <- relevel(x = cond, ref = "Group1")
SummarizedExperiment::colData(sca)$condition <- cond
fmla <- as.formula(
  object = paste0(" ~ ", paste(latent.vars.names, collapse = "+"))
)
zlmCond <- MAST::zlm(formula = fmla, sca = sca)
summaryCond <- MAST::summary(object = zlmCond, doLRT = 'conditionGroup2')
summaryDt <- summaryCond$datatable
summaryDt <- as.data.frame(summaryDt)
p_val <- summaryDt[summaryDt[, "component"] == "H", 4]
genes.return <- summaryDt[summaryDt[, "component"] == "H", 1]
to.return <- data.frame(p_val, row.names = genes.return)

de.result <- cbind(to.return, fc.results.1[rownames(x = to.return), , drop = FALSE])

de.result <- de.result[order(de.result$p_val, -abs(de.result$pct.1-de.result$pct.2)), ]
de.result$p_val_adj = p.adjust(
  p = de.result$p_val,
  method = "BH",
  n = nrow(x = de.result)
)
de.result
identical(de.result$p_val_adj, result1$p_val_adj)
```

```{r}
fmla
print(colnames(zlmCond@coefC))
```


```{r}
summaryDt[summaryDt[, "component"] == "H", ]
to.return
```



```{r}
summaryDt[summaryDt[, "component"] == "H", ]
summaryDt
table(summaryDt$component)
```
The hurdle model in MAST (Model-based Analysis of Single-cell Transcriptomics) is a statistical framework designed to account for the unique characteristics of single-cell RNA-seq data, particularly zero inflation (an excess of zero values) and variability in expression levels. It consists of two components: a logistic regression for the probability of expression and a linear regression for the magnitude of expression.

Key Features of the Hurdle Model in MAST
Zero Inflation:

Single-cell RNA-seq data often has many zero values due to technical dropouts or true biological absence of gene expression.
The hurdle model separates the analysis of whether a gene is expressed (non-zero) from how much it is expressed (continuous expression levels).
Two-Part Model:

The hurdle model combines:
Discrete Component (H Component):
A logistic regression that models the probability of a gene being expressed (non-zero values).
Predictors (e.g., condition, covariates) explain whether a gene is "on" or "off."
Continuous Component (C Component):
A linear regression that models the expression levels of a gene, conditional on it being expressed (non-zero values only).
Predictors explain the magnitude of expression.



H Component (Logistic Regression):
Significant terms indicate that a condition or covariate affects the probability of a gene being expressed.

C Component (Linear Regression):
Significant terms indicate that a condition or covariate affects the magnitude of gene expression among cells where it is expressed.

D Component (Combined Effect):
Combines the effects of detection and magnitude, providing an overall measure of differential expression.


### interaction


```{r}
logfc.threshold <- log(1)
latent.vars <- c('nCount_RNA','pmi','age_death', 'sex', 'apoe2', "cogdx2")
result1 <- FindMarkers(object = pbmc, slot = "data", assay = NULL, ident.1 = "F", ident.2 = "M", group.by = "sex", subset.ident = "T cells", logfc.threshold = logfc.threshold, test.use = 'MAST', latent.vars = latent.vars, min.pct = 0.1)
names(result1)[names(result1) == 'p_val_adj'] <- 'p_val_adj_bonferroni'
print(nrow(result1))
result1$p_val_adj <- p.adjust(result1$p_val, n=nrow(result1), method = "BH")
result1
result1[result1$p_val_adj < 0.05, ]

logfc.threshold <- log(1)
latent.vars <- c('nCount_RNA','pmi','age_death', 'apoe2', "cogdx2")
result1 <- FindMarkers(object = pbmc, slot = "data", assay = NULL, ident.1 = "F", ident.2 = "M", group.by = "sex", subset.ident = "T cells", logfc.threshold = logfc.threshold, test.use = 'MAST', latent.vars = latent.vars, min.pct = 0.1)
names(result1)[names(result1) == 'p_val_adj'] <- 'p_val_adj_bonferroni'
print(nrow(result1))
result1$p_val_adj <- p.adjust(result1$p_val, n=nrow(result1), method = "BH")
result1
result1[result1$p_val_adj < 0.05, ]
```

with Fold change
```{r}
pbmc.sub <- pbmc[, pbmc@meta.data$cell_type_high_resolution == "T cells"]
data.use.1 <- pbmc.sub[["RNA"]]
Idents(pbmc.sub) <- "sex"
latent.vars <- c('nCount_RNA','pmi','age_death', 'sex', 'apoe2', "cogdx2")

cellnames.use.1 <- colnames(data.use.1)

cells_1 <- my_IdentsToCells(object = pbmc.sub,ident.1 = "F",ident.2 = "M",cellnames.use = cellnames.use.1)

cells_1 <- sapply(X = cells_1,FUN = intersect,y = cellnames.use.1,simplify = FALSE,USE.NAMES = TRUE)

latent.vars.1 <- FetchData(object = pbmc.sub,vars = latent.vars,cells = c(cells_1$cells.1, cells_1$cells.2))

min.pct <- 0.1
min.diff.pct <- -Inf
slot <- "data"
only.pos <- FALSE
max.cells.per.ident <-  Inf
fc.results.1 <- FoldChange(object = data.use.1, slot = "data",
        cells.1 = cells_1$cells.1, cells.2 = cells_1$cells.2, features = NULL,
        pseudocount.use = 1, norm.method = "LogNormalize")

features <- NULL
features <- features %||% rownames(x = data.use.1@data)

# feature selection (based on percentages)
alpha.min <- pmax(fc.results.1$pct.1, fc.results.1$pct.2)
names(x = alpha.min) <- rownames(x = fc.results.1)
features <- names(x = which(x = alpha.min >= min.pct))
if (length(x = features) == 0) {
  warning("No features pass min.pct threshold; returning empty data.frame")
  return(fc.results.1[features, ])
}
alpha.diff <- alpha.min - pmin(fc.results.1$pct.1, fc.results.1$pct.2)
features <- names(
  x = which(x = alpha.min >= min.pct & alpha.diff >= min.diff.pct)
)
if (length(x = features) == 0) {
  warning("No features pass min.diff.pct threshold; returning empty data.frame")
  return(fc.results[features, ])
}

# feature selection (based on logFC)
if (slot != "scale.data") {
  total.diff <- fc.results.1[, 1] #first column is logFC
  names(total.diff) <- rownames(fc.results.1)
  features.diff <- if (only.pos) {
    names(x = which(x = total.diff >= logfc.threshold))
  } else {
    names(x = which(x = abs(x = total.diff) >= logfc.threshold))
  }
  features <- intersect(x = features, y = features.diff)
  if (length(x = features) == 0) {
    warning("No features pass logfc.threshold threshold; returning empty data.frame")
    return(fc.results.1[features, ])
  }
}

logfc.threshold <- log(1)
latent.vars <- latent.vars.1
cells.1 <- cells_1$cells.1
cells.2 <- cells_1$cells.2

data.use.2 <- data.use.1@data[features, c(cells.1, cells.2), drop = FALSE]

group.info <- data.frame(row.names = c(cells.1, cells.2))
latent.vars <- latent.vars %||% group.info
group.info[cells.1, "group"] <- "Group1"
group.info[cells.2, "group"] <- "Group2"
group.info[, "group"] <- factor(x = group.info[, "group"])
# group.info

latent.vars.names <- c("condition", colnames(x = latent.vars))
latent.vars <- cbind(latent.vars, group.info)
latent.vars$wellKey <- rownames(x = latent.vars)
# latent.vars

fdat <- data.frame(rownames(x = data.use.2))
colnames(x = fdat)[1] <- "primerid"
rownames(x = fdat) <- fdat[, 1]

sca <- MAST::FromMatrix(
  exprsArray = as.matrix(x = data.use.2),
  check_sanity = FALSE,
  cData = latent.vars,
  fData = fdat
)

cond <- factor(x = SummarizedExperiment::colData(sca)$group)
cond <- relevel(x = cond, ref = "Group1")
SummarizedExperiment::colData(sca)$condition <- cond
# fmla <- as.formula(
#   object = paste0(" ~ ", paste(latent.vars.names, collapse = "+"))
# )

sex_col <- factor(x = SummarizedExperiment::colData(sca)$sex)
sex_col <- relevel(x = sex_col, ref = "M")
SummarizedExperiment::colData(sca)$sex <- sex_col
apoe2_col <- factor(x = SummarizedExperiment::colData(sca)$apoe2)
apoe2_col <- relevel(x = apoe2_col, ref = "e33")
SummarizedExperiment::colData(sca)$apoe2 <- apoe2_col
cogdx2_col <- factor(x = SummarizedExperiment::colData(sca)$cogdx2)
cogdx2_col <- relevel(x = cogdx2_col, ref = "NCI")
SummarizedExperiment::colData(sca)$cogdx2 <- cogdx2_col
fmla <- as.formula(
  object = paste0(" ~ nCount_RNA + pmi + age_death + sex + apoe2 + cogdx2 + sex*apoe2*cogdx2")
)
fmla
zlmCond <- MAST::zlm(formula = fmla, sca = sca)
summaryCond <- MAST::summary(object = zlmCond, doLRT = 'sexF:apoe2e2x:cogdx2AD')
summaryDt <- summaryCond$datatable
summaryDt <- as.data.frame(summaryDt)
p_val <- summaryDt[summaryDt[, "component"] == "H", 4]
genes.return <- summaryDt[summaryDt[, "component"] == "H", 1]
to.return <- data.frame(p_val, row.names = genes.return)

de.result <- cbind(to.return, fc.results.1[rownames(x = to.return), , drop = FALSE])

de.result <- de.result[order(de.result$p_val, -abs(de.result$pct.1-de.result$pct.2)), ]
de.result$p_val_adj = p.adjust(
  p = de.result$p_val,
  method = "BH",
  n = nrow(x = de.result)
)
de.result
```

without foldchange - try to get coef
here i used F vs M but it does not matter. its the same as using AD vs NCI
```{r}
# pbmc <- ori.ls$Oligodendrocytes
# pbmc.sub <- pbmc[, pbmc@meta.data$cell_type_high_resolution == "Oli"]
# pbmc <- ori.ls$Excitatory_neurons_set3
# pbmc.sub <- pbmc[, pbmc@meta.data$cell_type_high_resolution == "Exc RELN CHD7"]
pbmc <- ori.ls$Immune_cells
pbmc.sub <- pbmc[, pbmc@meta.data$cell_type_high_resolution == "Mic P2RY12"]
data.use.1 <- pbmc.sub[["RNA"]]
Idents(pbmc.sub) <- "sex"
latent.vars <- c('nCount_RNA','pmi','age_death', 'sex', 'apoe2', "cogdx2")

cellnames.use.1 <- colnames(data.use.1)

cells_1 <- my_IdentsToCells(object = pbmc.sub,ident.1 = "F",ident.2 = "M",cellnames.use = cellnames.use.1)

cells_1 <- sapply(X = cells_1,FUN = intersect,y = cellnames.use.1,simplify = FALSE,USE.NAMES = TRUE)

latent.vars.1 <- FetchData(object = pbmc.sub,vars = latent.vars,cells = c(cells_1$cells.1, cells_1$cells.2))

min.pct <- 0.1
## not calculating FC (is not relevant, add coef result instead)
# min.diff.pct <- -Inf
# slot <- "data"
# only.pos <- FALSE
# max.cells.per.ident <-  Inf
# fc.results.1 <- FoldChange(object = data.use.1, slot = "data", 
#         cells.1 = cells_1$cells.1, cells.2 = cells_1$cells.2, features = NULL, 
#         pseudocount.use = 1, norm.method = "LogNormalize")
# 
# features <- NULL
# features <- features %||% rownames(x = data.use.1@data)
# 
# # feature selection (based on percentages)
# alpha.min <- pmax(fc.results.1$pct.1, fc.results.1$pct.2)
# names(x = alpha.min) <- rownames(x = fc.results.1)
# features <- names(x = which(x = alpha.min >= min.pct))
# if (length(x = features) == 0) {
#   warning("No features pass min.pct threshold; returning empty data.frame")
#   return(fc.results.1[features, ])
# }
# alpha.diff <- alpha.min - pmin(fc.results.1$pct.1, fc.results.1$pct.2)
# features <- names(
#   x = which(x = alpha.min >= min.pct & alpha.diff >= min.diff.pct)
# )
# if (length(x = features) == 0) {
#   warning("No features pass min.diff.pct threshold; returning empty data.frame")
#   return(fc.results[features, ])
# }
# 
# # feature selection (based on logFC)
# if (slot != "scale.data") {
#   total.diff <- fc.results.1[, 1] #first column is logFC
#   names(total.diff) <- rownames(fc.results.1)
#   features.diff <- if (only.pos) {
#     names(x = which(x = total.diff >= logfc.threshold))
#   } else {
#     names(x = which(x = abs(x = total.diff) >= logfc.threshold))
#   }
#   features <- intersect(x = features, y = features.diff)
#   if (length(x = features) == 0) {
#     warning("No features pass logfc.threshold threshold; returning empty data.frame")
#     return(fc.results.1[features, ])
#   }
# }
r0.df <- as.data.frame(t(data.use.1@data))
keep <- which(colSums(r0.df) != 0 )
r0.df <- r0.df[, keep]
dim(r0.df)
r0.df$sex_apoe2_cogdx <- pbmc@meta.data$sex_apoe2_cogdx[match(rownames(r0.df), rownames(pbmc@meta.data))]
table(r0.df$sex_apoe2_cogdx)
# genes_to_keep <- r0.df %>%
#   dplyr::select(-sex_apoe2_cogdx) %>%  # Select all gene columns
#   dplyr::mutate(sex_apoe2_cogdx = r0.df$sex_apoe2_cogdx) %>%  # Add sex_apoe2_cogdx back for context
#   tidyr::pivot_longer(
#     cols = -sex_apoe2_cogdx,  # Reshape gene columns to long format
#     names_to = "gene",
#     values_to = "expression"
#   ) %>%
#   dplyr::filter(expression != 0) %>%  # Keep only non-zero expression values
#   dplyr::group_by(gene, sex_apoe2_cogdx) %>%  # Group by gene and sex_apoe2_cogdx
#   dplyr::summarise(
#     count = dplyr::n(),  # Count non-zero samples per group
#     total = sum(r0.df$sex_apoe2_cogdx == sex_apoe2_cogdx[1]),  # Total samples in each group
#     .groups = "drop"
#   ) %>%
#   dplyr::mutate(prop = count / total) %>%  # Calculate proportion of non-zero samples
#   dplyr::group_by(gene) %>%  # Group by gene for the next step
#   dplyr::filter(any(prop >= min.pct)) %>%  # Check if any group has proportion >= 0.1
#   dplyr::ungroup() %>%  # Ungroup for final distinct operation
#   dplyr::distinct(gene) %>%  # Extract unique genes
#   dplyr::pull(gene)  # Get the list of gene names

process_chunk <- function(chunk) {
  chunk %>%
    tidyr::pivot_longer(
      cols = -sex_apoe2_cogdx,
      names_to = "gene",
      values_to = "expression"
    ) %>%
    dplyr::filter(expression != 0) %>%
    dplyr::group_by(gene, sex_apoe2_cogdx) %>%
    dplyr::summarise(
      count = dplyr::n(),
      total = sum(chunk$sex_apoe2_cogdx == sex_apoe2_cogdx[1]),
      .groups = "drop"
    ) %>%
    dplyr::mutate(prop = count / total) %>%
    dplyr::group_by(gene) %>%
    dplyr::filter(any(prop >= 0.1)) %>%
    dplyr::ungroup() %>%
    dplyr::distinct(gene) %>%
    dplyr::pull(gene)
}

# Split the data into chunks
chunk_size <- 100  # Adjust the chunk size based on your memory limits
gene_chunks <- split(names(r0.df)[!names(r0.df) %in% "sex_apoe2_cogdx"], 
                     ceiling(seq_along(names(r0.df)[!names(r0.df) %in% "sex_apoe2_cogdx"]) / chunk_size))

# Process each chunk and combine the results
genes_to_keep <- lapply(gene_chunks, function(chunk) {
  process_chunk(r0.df[, c(chunk, "sex_apoe2_cogdx")])
}) %>%
  unlist() %>%
  unique()

length(genes_to_keep)
features <- genes_to_keep


logfc.threshold <- log(1)
latent.vars <- latent.vars.1
cells.1 <- cells_1$cells.1
cells.2 <- cells_1$cells.2

data.use.2 <- data.use.1@data[features, c(cells.1, cells.2), drop = FALSE]

group.info <- data.frame(row.names = c(cells.1, cells.2))
latent.vars <- latent.vars %||% group.info
group.info[cells.1, "group"] <- "Group1"
group.info[cells.2, "group"] <- "Group2"
group.info[, "group"] <- factor(x = group.info[, "group"])
# group.info

latent.vars.names <- c("condition", colnames(x = latent.vars))
latent.vars <- cbind(latent.vars, group.info)
latent.vars$wellKey <- rownames(x = latent.vars)
# latent.vars

fdat <- data.frame(rownames(x = data.use.2))
colnames(x = fdat)[1] <- "primerid"
rownames(x = fdat) <- fdat[, 1]

sca <- MAST::FromMatrix(
  exprsArray = as.matrix(x = data.use.2),
  check_sanity = FALSE,
  cData = latent.vars,
  fData = fdat
)

cond <- factor(x = SummarizedExperiment::colData(sca)$group)
cond <- relevel(x = cond, ref = "Group1")
SummarizedExperiment::colData(sca)$condition <- cond
# fmla <- as.formula(
#   object = paste0(" ~ ", paste(latent.vars.names, collapse = "+"))
# )

sex_col <- factor(x = SummarizedExperiment::colData(sca)$sex)
sex_col <- relevel(x = sex_col, ref = "M")
SummarizedExperiment::colData(sca)$sex <- sex_col
apoe2_col <- factor(x = SummarizedExperiment::colData(sca)$apoe2)
apoe2_col <- relevel(x = apoe2_col, ref = "e33")
SummarizedExperiment::colData(sca)$apoe2 <- apoe2_col
cogdx2_col <- factor(x = SummarizedExperiment::colData(sca)$cogdx2)
cogdx2_col <- relevel(x = cogdx2_col, ref = "NCI")
SummarizedExperiment::colData(sca)$cogdx2 <- cogdx2_col
fmla <- as.formula(
  object = paste0(" ~ nCount_RNA + pmi + age_death + sex + apoe2 + cogdx2 + sex*apoe2*cogdx2")
)
fmla
zlmCond <- MAST::zlm(formula = fmla, sca = sca)
summaryCond <- MAST::summary(object = zlmCond, doLRT = 'sexF:apoe2e2x:cogdx2AD')
# summaryCond <- MAST::summary(object = zlmCond, doLRT = 'cogdx2AD')
summaryDt <- summaryCond$datatable
summaryDt <- as.data.frame(summaryDt)
p_val <- summaryDt[summaryDt[, "component"] == "H", 4]
coef <- summaryDt[summaryDt[, "component"] == "H", 7]
genes.return <- summaryDt[summaryDt[, "component"] == "H", 1]
to.return <- data.frame(p_val, coef, row.names = genes.return)

de.result <- to.return

de.result <- de.result[order(de.result$p_val), ]
de.result$p_val_adj = p.adjust(
  p = de.result$p_val,
  method = "BH",
  n = nrow(x = de.result)
)
de.result
```
Here, doLRT = "sex:apoe:condition" performs a likelihood ratio test for the interaction term across all genes 
Interpret the Results:
Genes with significant p-values for the interaction term (sex:condition) are those where the effect of sex on expression levels depends on the condition (or vice versa).
This means that for these genes, there’s a significant difference in how sex affects gene expression in one condition compared to the other.



new: doLRT = 'cogdx2AD'
even with not complicated term, H still has no coef
```{r}
table(summaryDt$component)
for (conp in unique(summaryDt$component)){
  print(summaryDt[summaryDt$component == conp, ])
}
```


```{r}
de.result[de.result$p_val_adj < 0.05, ]
```


```{r}
sca@colData
print(colnames(zlmCond@coefC))
```




```{r}
r0.df <- as.data.frame(t(data.use.1@data))
keep <- which(colSums(r0.df) != 0 )
r0.df <- r0.df[, keep]
dim(r0.df)
r0.df$sex_apoe2_cogdx <- pbmc@meta.data$sex_apoe2_cogdx[match(rownames(r0.df), rownames(pbmc@meta.data))]
table(r0.df$sex_apoe2_cogdx)
genes_to_keep <- r0.df %>%
  dplyr::select(-sex_apoe2_cogdx) %>%  # Select all gene columns
  dplyr::mutate(sex_apoe2_cogdx = r0.df$sex_apoe2_cogdx) %>%  # Add sex_apoe2_cogdx back for context
  tidyr::pivot_longer(
    cols = -sex_apoe2_cogdx,  # Reshape gene columns to long format
    names_to = "gene",
    values_to = "expression"
  ) %>%
  dplyr::filter(expression != 0) %>%  # Keep only non-zero expression values
  dplyr::group_by(gene, sex_apoe2_cogdx) %>%  # Group by gene and sex_apoe2_cogdx
  dplyr::summarise(
    count = dplyr::n(),  # Count non-zero samples per group
    total = sum(r0.df$sex_apoe2_cogdx == sex_apoe2_cogdx[1]),  # Total samples in each group
    .groups = "drop"
  ) %>%
  dplyr::mutate(prop = count / total) %>%  # Calculate proportion of non-zero samples
  dplyr::group_by(gene) %>%  # Group by gene for the next step
  dplyr::filter(any(prop >= 0.1)) %>%  # Check if any group has proportion >= 0.1
  dplyr::ungroup() %>%  # Ungroup for final distinct operation
  dplyr::distinct(gene) %>%  # Extract unique genes
  dplyr::pull(gene)  # Get the list of gene names

length(genes_to_keep)
setdiff(names(keep), genes_to_keep)[1:10]
```

```{r}
setdiff(names(keep), genes_to_keep)[1:10]
```


```{r}
non_zero_samples <- r0.df[r0.df[["FAM41C"]] != 0, ]
dim(non_zero_samples)
group_counts <- table(non_zero_samples$sex_apoe2_cogdx)
group_counts <- group_counts / sapply(names(group_counts), function(x) sum(r0.df$sex_apoe2_cogdx == x))

group_counts
```




```{r}
library(SeuratDisk)
seuratObject <- LoadH5Seurat("/sc/arion/projects/adineto/sea_ad/Data/snRNAseq/PFC/SEAAD_A9_RNAseq_final-nuclei.2024-02-13.h5ad")
```



# After analysis:

```{r}
all.res <- NULL

for (c in processed.ls){
  if(!file.exists(paste0("/sc/arion/projects/zhangb03a/yug03/mathys/snRNA/interaction_regression/MAST/", c, ".tsv"))){
    print(which(processed.ls == c))
    print(c)
  }
  else{
    f <- read.csv(paste0("/sc/arion/projects/zhangb03a/yug03/mathys/snRNA/interaction_regression/MAST/", c, ".tsv"), sep="\t")
    all.res <- rbind(all.res, f)
  }
  
}
# print(file)
cat("\n")
  

all.res
```

```{r}
all.res$Symbol_Celltype <- paste0(all.res$Symbol, "_", all.res$Celltype)

all.res.e2x <- all.res[all.res$Contrast == "sexF:apoe2e2x:cogdx2AD", ]
all.res.e2x
all.res.e4x <- all.res[all.res$Contrast == "sexF:apoe2e4x:cogdx2AD", ]
all.res.e4x

all.res.e2x.sig <- all.res.e2x[all.res.e2x$p_val_adj < 0.05, ]
all.res.e2x.sig
length(unique(all.res.e2x.sig$Symbol))
all.res.e4x.sig <- all.res.e4x[all.res.e4x$p_val_adj < 0.05, ]
all.res.e4x.sig
length(unique(all.res.e4x.sig$Symbol))
```

```{r}
table(all.res.e2x.sig$Celltype)
all.res.e4x.sig[all.res.e4x.sig$Symbol == "CLU", ]
```

```{r, fig.height=7.5, fig.width=12}
all.res.e2x.sig

celltype_counts.e2x <- all.res.e2x.sig %>%
  group_by(Celltype) %>%
  summarise(SymbolCount = n(), .groups = "drop") %>%
  right_join(celltype.df, by = c("Celltype" = "subcluster")) %>%  # Join with all cell types from `celltype.df`
  mutate(SymbolCount = ifelse(is.na(SymbolCount), 0, SymbolCount)) 
celltype_counts.e2x$C <- celltype.df$celltype[match(celltype_counts.e2x$Celltype, celltype.df$subcluster)]
celltype_counts.e2x

p1 <- ggplot(celltype_counts.e2x, aes(x = Celltype, y = SymbolCount, fill = C)) +
  geom_bar(stat = "identity") + theme_minimal() + facet_grid(.~C, scales="free", space = "free") + 
  labs(
    title = "sexF:apoee2x:cogdxAD", x = "", y = "# Significant Genes") +
  theme(legend.position = "none", axis.text.x = element_blank()) 


all.res.e4x.sig

celltype_counts.e4x <- all.res.e4x.sig %>%
 group_by(Celltype) %>%
  summarise(SymbolCount = n(), .groups = "drop") %>%
  right_join(celltype.df, by = c("Celltype" = "subcluster")) %>%  # Join with all cell types from `celltype.df`
  mutate(SymbolCount = ifelse(is.na(SymbolCount), 0, SymbolCount)) 
celltype_counts.e4x$C <- celltype.df$celltype[match(celltype_counts.e4x$Celltype, celltype.df$subcluster)]
celltype_counts.e4x

p2 <- ggplot(celltype_counts.e4x, aes(x = Celltype, y = SymbolCount, fill = C)) +
  geom_bar(stat = "identity") + theme_minimal() + facet_grid(.~C, scales="free", space = "free") + 
  labs(
    title = "sexF:apoee4x:cogdxAD", x = "", y = "# Significant Genes") +
  theme(legend.position = "none", axis.text.x = element_text(angle = 90, hjust = 1)) 

p1 / p2 + plot_layout(heights = c(1, 1)) 
```
```{r}
celltype.df
```




maybe use C / D results to further filter sig genes? 
```{r}
all.summaryDt <- NULL

for (c in processed.ls){
  if(!file.exists(paste0("/sc/arion/projects/zhangb03a/yug03/mathys/snRNA/interaction_regression/MAST/", c, "_summaryDt.RDS"))){
    print(which(processed.ls == c))
    print(c)
  }
  else{
    f <- readRDS(paste0("/sc/arion/projects/zhangb03a/yug03/mathys/snRNA/interaction_regression/MAST/", c, "_summaryDt.RDS"))
    all.summaryDt <- rbind(all.summaryDt, f)
  }
  
}
# print(file)
cat("\n")
  

all.summaryDt
```

```{r}
# all.summaryDt$p_val_adj <- NA
# for(c in unique(all.summaryDt$Celltype)){
#   for (contrast in c("sexF:apoe2e2x:cogdx2AD", "sexF:apoe2e4x:cogdx2AD")){
#     for (comp in unique(all.summaryDt$component)){
#       all.summaryDt[all.summaryDt$Celltype == c & 
#                     all.summaryDt$contrast == contrast & 
#                     all.summaryDt$component == comp, "p_val_adj"] <- p.adjust(all.summaryDt[all.summaryDt$Celltype == c &
#                                                                                            all.summaryDt$contrast == contrast &
#                                                                                           all.summaryDt$component == comp, ]$`Pr(>Chisq)`, method = "BH")
#     }
#   }
# }
# sum(is.na(all.summaryDt$p_val_adj))
# all.summaryDt
# 
# saveRDS(all.summaryDt, file = paste0("/sc/arion/projects/zhangb03a/yug03/mathys/snRNA/interaction_regression/MAST/all_summaryDt_adjp.RDS"))
```

```{r}
all.summaryDt <- readRDS("/sc/arion/projects/zhangb03a/yug03/mathys/snRNA/interaction_regression/MAST/all_summaryDt_adjp.RDS")
all.summaryDt
```


```{r}
table(all.summaryDt$component)
all.summaryDt[all.summaryDt$component == "H" & all.summaryDt$primerid== "MRO", ]
```


```{r}
all.summaryDt$Symbol_Celltype <- paste0(all.summaryDt$primerid, "_", all.summaryDt$Celltype)

all.summaryDt.e2x <- all.summaryDt[all.summaryDt$component == "H" & all.summaryDt$contrast == "sexF:apoe2e2x:cogdx2AD", ]
all.summaryDt.e2x
all.summaryDt.e4x <- all.summaryDt[all.summaryDt$component == "H" & all.summaryDt$contrast == "sexF:apoe2e4x:cogdx2AD", ]
all.summaryDt.e4x

all.summaryDt.e2x.sig <- all.summaryDt.e2x[all.summaryDt.e2x$p_val_adj < 0.05, ]
all.summaryDt.e2x.sig
length(unique(all.summaryDt.e2x.sig$primerid))
all.summaryDt.e4x.sig <- all.summaryDt.e4x[all.summaryDt.e4x$p_val_adj < 0.05, ]
all.summaryDt.e4x.sig
length(unique(all.summaryDt.e4x.sig$primerid))
```


(MAST::summary(object = zlmCond, doLRT = 'sexF:apoe2e2x:cogdx2AD') also have "sexF:apoe2e4x:cogdx2AD" term but has no p value
MAST::summary(object = zlmCond, doLRT = 'sexF:apoe2e4x:cogdx2AD') also have "sexF:apoe2e2x:cogdx2AD" term but has no p value 
so I need to exclude the redundent results)
```{r}
all.summaryDt.e2x.c <- all.summaryDt[all.summaryDt$component == "C" & all.summaryDt$contrast == "sexF:apoe2e2x:cogdx2AD" & !is.na(all.summaryDt$`Pr(>Chisq)`), ]
all.summaryDt.e2x.c
all.summaryDt.e4x.c <- all.summaryDt[all.summaryDt$component == "C" & all.summaryDt$contrast == "sexF:apoe2e4x:cogdx2AD"& !is.na(all.summaryDt$`Pr(>Chisq)`), ]
all.summaryDt.e4x.c

# all.summaryDt.e2x.c.sig <- all.summaryDt.e2x.c[all.summaryDt.e2x.c$p_val_adj < 0.05 , ]
all.summaryDt.e2x.c.sig <- all.summaryDt.e2x.c[all.summaryDt.e2x.c$p_val_adj < 0.05 & abs(all.summaryDt.e2x.c$coef) > 0.2, ]
length(unique(all.summaryDt.e2x.c.sig$primerid))
all.summaryDt.e2x.c.sig

# all.summaryDt.e4x.c.sig <- all.summaryDt.e4x.c[all.summaryDt.e4x.c$p_val_adj < 0.05 , ]
all.summaryDt.e4x.c.sig <- all.summaryDt.e4x.c[all.summaryDt.e4x.c$p_val_adj < 0.05 & abs(all.summaryDt.e2x.c$coef) > 0.2, ]
length(unique(all.summaryDt.e4x.c.sig$primerid))
all.summaryDt.e4x.c.sig
```


```{r}
all.summaryDt.e2x.d <- all.summaryDt[all.summaryDt$component == "D" & all.summaryDt$contrast == "sexF:apoe2e2x:cogdx2AD" & !is.na(all.summaryDt$`Pr(>Chisq)`), ]
all.summaryDt.e2x.d
all.summaryDt.e4x.d <- all.summaryDt[all.summaryDt$component == "D" & all.summaryDt$contrast == "sexF:apoe2e4x:cogdx2AD"& !is.na(all.summaryDt$`Pr(>Chisq)`), ]
all.summaryDt.e4x.d

all.summaryDt.e2x.d.sig <- all.summaryDt.e2x.d[all.summaryDt.e2x.d$p_val_adj < 0.05, ]
length(unique(all.summaryDt.e2x.d.sig$primerid))
all.summaryDt.e2x.d.sig

all.summaryDt.e4x.d.sig <- all.summaryDt.e4x.d[all.summaryDt.e4x.d$p_val_adj < 0.05, ]
length(unique(all.summaryDt.e4x.d.sig$primerid))
all.summaryDt.e4x.d.sig
```

```{r}
e2x.cd.sig <- intersect(all.summaryDt.e2x.c.sig$Symbol_Celltype, all.summaryDt.e2x.d.sig$Symbol_Celltype)
e4x.cd.sig <- intersect(all.summaryDt.e4x.c.sig$Symbol_Celltype, all.summaryDt.e4x.d.sig$Symbol_Celltype)


all.summaryDt.e2x.sig.cd <- all.summaryDt.e2x.sig[all.summaryDt.e2x.sig$Symbol_Celltype %in% e2x.cd.sig, ]
table(all.summaryDt.e2x.sig.cd$Celltype)
all.summaryDt.e4x.sig.cd <- all.summaryDt.e4x.sig[all.summaryDt.e4x.sig$Symbol_Celltype %in% e2x.cd.sig, ]
table(all.summaryDt.e4x.sig.cd$Celltype)
```

```{r}
r <- all.summaryDt.e4x.c.sig[all.summaryDt.e4x.c.sig$Symbol_Celltype %in% e4x.cd.sig, ]
r[order(abs(r$coef), decreasing = T), ]
intersect(r$primerid, risk_genes)
```


```{r}
length(intersect(all.summaryDt.e2x.sig$Symbol_Celltype, all.summaryDt.e4x.sig$Symbol_Celltype))
length(intersect(all.summaryDt.e2x.c.sig$Symbol_Celltype, all.summaryDt.e4x.c.sig$Symbol_Celltype))
length(intersect(intersect(all.summaryDt.e2x.sig$Symbol_Celltype, all.summaryDt.e4x.sig$Symbol_Celltype), intersect(all.summaryDt.e2x.c.sig$Symbol_Celltype, all.summaryDt.e4x.c.sig$Symbol_Celltype)))

r <- all.summaryDt.e4x[all.summaryDt.e4x$Symbol_Celltype %in% intersect(intersect(all.summaryDt.e2x.sig$Symbol_Celltype, all.summaryDt.e4x.sig$Symbol_Celltype), intersect(all.summaryDt.e2x.c.sig$Symbol_Celltype, all.summaryDt.e4x.c.sig$Symbol_Celltype)), ]
r[order(r$p_val_adj), ]
```

```{r}
all.summaryDt.e4x.sig <- all.summaryDt.e4x[all.summaryDt.e4x$p_val_adj < 0.01, ]
# all.summaryDt.e4x.sig
length(unique(all.summaryDt.e4x.sig$primerid))

all.summaryDt.e4x.sig[order(all.summaryDt.e4x.sig$p_val_adj, decreasing = T), ]
# table(all.summaryDt.e4x.sig$Celltype)

all.summaryDt.e4x.c.sig[all.summaryDt.e4x.c.sig$Symbol_Celltype == "AC105942.1_Ast GRM3", ]
```


```{r, fig.width=7, fig.height=5}
gene_interaction <- function(gene, cluster){
  cellfile <- celltype.df[celltype.df$subcluster == cluster, "processed.ls3"]
  pbmc <- ori.ls[[cellfile]]
  pbmc.sub <- pbmc[, pbmc@meta.data$cell_type_high_resolution == cluster]
  
  v <- VlnPlot(pbmc.sub, features = gene, group.by = "sex_apoe2", split.by = "cogdx2", pt.size = 0) + geom_boxplot(width=0.1, position=position_dodge(.9)) + ggtitle(paste(cluster, gene))
  print(v)
}
gene_interaction("HSPH1", "Mic P2RY12")
gene_interaction("CLU", "Exc L2-3 CBLN2 LINC02306")
gene_interaction("HSPH1", "Exc L2-3 CBLN2 LINC02306")
gene_interaction("AC105942.1", "Ast GRM3")
```








```{r gene_FC_heatmap}
all.res.merged <- do.call(rbind, diff.con.cogdx.AD.NCI.s5.res)

gene_FC_heatmap <- function(all.res, the.gene, c){
  gene.df <- as.data.frame(matrix(nrow = 6, ncol = length(unique(grep(c, all.meta.data$cell_type_high_resolution, value = T)))))
  gene.df.anno <- as.data.frame(matrix(nrow = 6, ncol = length(unique(grep(c, all.meta.data$cell_type_high_resolution, value = T)))))
  rownames(gene.df) <- c("F_e2x", "F_e33", "F_e4x", "M_e2x", "M_e33", "M_e4x")
  rownames(gene.df.anno) <- c("F_e2x", "F_e33", "F_e4x", "M_e2x", "M_e33", "M_e4x")
  colnames(gene.df) <- unique(grep(c, all.meta.data$cell_type_high_resolution, value = T))
  colnames(gene.df.anno) <- unique(grep(c, all.meta.data$cell_type_high_resolution, value = T))
  
  for (i in 1:nrow(gene.df)){
    contrast <- paste0(rownames(gene.df)[i], "_AD_vs_", rownames(gene.df)[i], "_NCI" )
    for (cluster in colnames(gene.df)){
      if (nrow(all.res[all.res$Contrast == contrast & all.res$Symbol == the.gene & all.res$Celltype == cluster, ]) > 0){
        gene.df[i, cluster] <- all.res[all.res$Contrast == contrast & all.res$Symbol == the.gene & all.res$Celltype == cluster, "avg_log2FC"]
        adjp <- all.res[all.res$Contrast == contrast & all.res$Symbol == the.gene & all.res$Celltype == cluster, "p_val_adj"]
        gene.df.anno[i, cluster] <- ifelse(adjp < 0.05 & abs(gene.df[i, cluster]) > log2(1.3), TRUE, FALSE)
      }
      else{
        gene.df[i, cluster] <- 0
        gene.df.anno[i, cluster] <- FALSE
      }
    }
  }
  gene.df <- as.matrix(gene.df)
  h <- Heatmap(gene.df,  row_names_side = "left", cluster_rows = T, cluster_columns = T, column_title=paste(the.gene, c), 
               heatmap_legend_param = list(title = "log2FC", title_gp = gpar(fontsize = 15, fontface = "plain"),title_position = "leftcenter-rot",labels_gp = gpar(fontsize = 15)), 
               cell_fun = function(j, i, x, y, w, h, fill) {
                 if(gene.df.anno[i, j] == TRUE) { grid.text("*", x, y) }  }
               )
  print(h) 
}
```

```{r}
gene_FC_heatmap(all.res.merged, "CHD7", "Exc RELN CHD7")
```


```{r, fig.height=7.5, fig.width=12}
all.summaryDt.e2x.sig.cd

celltype_counts.e2x <- all.summaryDt.e2x.sig.cd %>%
  group_by(Celltype) %>%
  summarise(SymbolCount = n(), .groups = "drop") %>%
  right_join(celltype.df, by = c("Celltype" = "subcluster")) %>%  # Join with all cell types from `celltype.df`
  mutate(SymbolCount = ifelse(is.na(SymbolCount), 0, SymbolCount)) 
celltype_counts.e2x$C <- celltype.df$celltype[match(celltype_counts.e2x$Celltype, celltype.df$subcluster)]
celltype_counts.e2x

p1 <- ggplot(celltype_counts.e2x, aes(x = Celltype, y = SymbolCount, fill = C)) +
  geom_bar(stat = "identity") + theme_minimal() + facet_grid(.~C, scales="free", space = "free") + 
  labs(
    title = "sexF:apoee2x:cogdxAD", x = "", y = "# Significant Genes") +
  theme(legend.position = "none", axis.text.x = element_blank()) 


all.summaryDt.e4x.sig.cd

celltype_counts.e4x <- all.summaryDt.e4x.sig.cd %>%
 group_by(Celltype) %>%
  summarise(SymbolCount = n(), .groups = "drop") %>%
  right_join(celltype.df, by = c("Celltype" = "subcluster")) %>%  # Join with all cell types from `celltype.df`
  mutate(SymbolCount = ifelse(is.na(SymbolCount), 0, SymbolCount)) 
celltype_counts.e4x$C <- celltype.df$celltype[match(celltype_counts.e4x$Celltype, celltype.df$subcluster)]
celltype_counts.e4x

p2 <- ggplot(celltype_counts.e4x, aes(x = Celltype, y = SymbolCount, fill = C)) +
  geom_bar(stat = "identity") + theme_minimal() + facet_grid(.~C, scales="free", space = "free") + 
  labs(
    title = "sexF:apoee4x:cogdxAD", x = "", y = "# Significant Genes") +
  theme(legend.position = "none", axis.text.x = element_text(angle = 90, hjust = 1)) 

p1 / p2 + plot_layout(heights = c(1, 1)) 
```


